Discovery of Rotational Modulations in the Planetary-mass Companion 2M1207b: A Slow Rotation Period and Cloud Height in a Low Gravity Atmosphere

### 2015-09-25 summary/letter/letter_vAD.tex
submitted

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{natbib}
\bibliographystyle{apj}
\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=magenta]{hyperref}
\shortauthors{Zhou et al.}

@@

\title{Discovery of Rotational Modulations in the Planetary-mass
  Companion 2M1207b: A Slow Rotation Period and Cloud Height in a Low
  Gravity Atmosphere}
\shorttitle{Variability of 2M1207 B}
\author{Yifan Zhou, Daniel Apai, Glenn Schneider}
\affil{Steward Observatory, the University of Arizona}
\author{Mark Marley}
\affil{NASA Ames Research Center}

\begin{abstract}
  Rotational modulations of brown dwarfs have recently provided
  powerful constraints on the properties of ultra-cool atmospheres,
  including longitudinal and vertical cloud structures and cloud
  evolution. Furthermore, detection of periodic light curve variations
  can directly probe the rotational periods of ultra-cool objects.

  We present here, for the first time, time-resolved high-precision
  photometric measurements of a planetary-mass companion, 2MASS1207b,
  to a brown dwarf primary. Using HST/WFC3 and point spread
  function %in
  combination with two spacecraft roll angles we detect photometric
  modulations in the light curve. The amplitude is 0.9\% in the F160W
  and 1.5\% in the F125W filters; we find a consistent period and
  similar phase in both bands. Joint fit to the light curve in both
  bands suggest a period of $10.2^{+0.9}_{-0.8}$ h. The relative
  amplitudes in the two filters are very similar to that found in a
  recent study of a field (high-gravity) L-dwarf, suggesting that the
  cloud structures that introduce the photometric modulations are
  similar in high- and low-gravity objects. Importantly, our study
  also measures, for the first time, the rotational period for
  directly imaged planetary-mass companion.
\end{abstract}

\keywords{brown dwarfs -- planets and satellites: atmospheres -- planets
  and satellites: individual (2M1207b) -- techniques: photometric}
\maketitle
%
\section{Introduction}

% Cloud properties in high- and low-gravity objects -- thick clouds: an
% explanation for very red and faint near-infrared fluxes of directly
% imaged planets
Presence of condensate clouds is one the most unique and essential
features of the ultra-cool atmosphere of direct imaged exoplanet and brown
dwarfs. The formation and properties of condensate clouds of are
intensively studied in recent years \citep[e.g.][]{Ackerman2001,
  Helling2008, Allard2012}, and great improvement has been
achieved in understanding the cloud behavior across different spectral
types, especially in the explanation of L-T transition \citep{Marley2010}.
Surface gravity is suggested to be another key parameters 
in defining cloud structures. Low surface gravity objects (e.g. HR8799 bcd, 2M1207
b) have significantly redder color compared to their spectral type
matched brown dwarfs, which are best explained by the existence of 
thick clouds \citep{Barman2011, Skemer2011, Skemer2012}. However, due to lack of observational constraint, the
behavior of clouds in the atmospheres of low surface gravity objects
are not as well as modeled as high surface gravity counterpart.

Intensity modulations introduced by heterogeneous clouds can be
directly observed and studied via time resolved observation and
rotational mapping. These techniques isolate the effect of cloud
properties and obtained great success in determining the rotation
period and unveiling the structures of the atmosphere of brown dwarfs
\citep[e.g.][]{Apai2013,Buenzli2012,Buenzli2015,Burgasser2013,Radigan2012,Yang2014,Metchev2015}. \cite{Kostov2013}
demonstrated the great potential of these techniques in the study of the cloud properties of
direct imaged exoplanets and planetary mass companions. However, high
brightness contrast makes it much more challenging for direct
imaged exoplanet and planetary mass companions than brown
dwarfs to acquire high-precision light curves.

2M1207 b \citep{Chauvin2005} is a planetary mass companion that has mass of
$\sim 4M_{\mathrm{Jup}}$ and an angular separation of $0.78''$ to its host
2M1207 A,
an M8 brown dwarf. Relative moderate contrast (6.7 mag in J band, \cite{Mohanty2007}) makes it one
the most feasible planetary mass companions for high-precision
time-resolved observation and the best starting points for developing
strategies for time-resolved observation and data reduction in high
contrast case.

@@

We obtained direct images of the 2M1207A+b
system on UT 2014 April 11from 08:07:47 to 16:53:18 using HST and its Wide Field Camera 3 \citep[WFC3,
][]{Kimble2008} in the frame of the HST Proposal GO-13418 (PI:
D. Apai). We acquired the observations in filters F125W
($\lambda_{\mbox{pivot}}$ = 1245.9 nm, full width at half maximum
(FWHM) = 301.5 nm) and F160W ($\lambda_{\mbox{pivot}}$ 1540.52, FWHM =
287.9 nm), roughly corresponding to the J and H bands. The WFC3 pixel
scale is $\approx$13mas. We used the $256\times256$ pixels sub-array
mode to avoid memory dumps during the observations.  In order to
provide a near-continuous coverage for detecting modulations we
observed the 2M1207 system in six consecutive HST orbits, obtaining
data with cadence of $\sim1.5$ minutes over a baseline of 8 hours and
40 minutes. The observations were interrupted by 58 minutes long Earth
occultations every 94 minutes.

The observations applied space craft rolls between each two orbits to
allow roll-subtraction of the primary \citep[e.g.][]{Song2006}. The
telescope roll angles for orbits 1, 3, and 5, and those taken in
orbits 2, 4, and 6 differed by $25^{\circ}$. At the separation of
2M1207b this angle difference corresponds to a displacement of
$0.34''$ or 2.75 and 2.30 resolution elements in F125W and F160W,
respectively. In each orbit we took thirteen SPARS10 non-destructive
read-outs with NSAMP=10, alternating between F160W and F125W filters,
with 2--3 identical exposures in one exposure sequence. To improve
sampling and reduce the risk that point spread funciton (PSF) lays on
bad pixels, we applied a 4-point dither pattern with differential "X/Y"
offsets of 1.375" in the detector frame, providing optimal
non-integral (half pixel) stepping of 10.5 and 8.5 pixels in F125W and
F160W, respectively. Over the 6 orbits, we obtained 70 images with 10
non-destructive read-outs in F125W and 64 images in F160W with
exposure time of 88.4~s in each filters.

@@

%Although \cite{Mandell2013} stated that WFC3 IR time series
%extract from {\flt} have a rms 1.3 times larger than that obtained
%from {\ima}, \flt keeps all pixels of the image of 2M1207 A
%unsaturated, while in 80\% of non-destructive reads the cores of
%2M1207 A are saturated. 
%Thus \flt help better align the images for
%primary subtraction, and separate the flux of 2M1207 B from that of
%2M1207A.

%The small angular separation of 2M1207 A and B (as shown in Figure
%\ref{fig:1}) makes precise primary star subtraction and photometry
%very difficult. 
%On the under-sampled WFC3 IR detector (plate scale
  %$\sim 0.13''\mbox{pixel}^{-1}$ \cite{dressel2012wide}), the primary
%and the secondary only separate by $\sim 6$ pixels, which is about 5
%times of the FWHM of the PSF. In addition, under-sampling causes
%significant artifacts when shifting the images for registration no
%matter what interpolation method is used.
 \begin{figure*}
  \centering
  \plotone{systematics}
  \caption{F125W (left) and F160W (right) light curves under different
    variability verification tests. Individual measurements are
    plotted with gray crosses. Photometric measurements of the same exposure
    sequence are binned, and binned photometry are plotted with points
    or squares. Best fitted sinusoidal curves are plotted with solid
    lines. {\em Upper}: binned measurements taken in dithering
    position 1 and 3 (red points) and that taken in 2 and 4 (blue
    squares) are plotted with different symbols. They demonstrate same
    trend of variation. In upper left panel, green line is sinusoidal
    wave fitted with all parameter set free, and purple line is
    sinusoidal wave fitted with period set the same as that of
    F125W. {\em Middle}: sinusoidal waves fitted without using the
    data taken in Orbit \#1. These curves are almost identical to the
    curves plotted in upper panel. {\em Lower}: photometry measured
    with AFEM-added images and best fitted sinusoidal curves. These
    points and curves are also almost identical to those plotted
    in the upper panel.}
  \label{fig:2}
\end{figure*}

@@

We started the reduction from the the \flt files produced by the
WFC3's \texttt{calwfc3} pipeline. We did not opt to use the more
advance \ima pipeline products because these provided less information
on 2M1207A, which saturated after the first few read-outs.  The \flt
are results of basic calibration, including dark current correction,
non-linearity correction, flat field correction, as well as
up-the-ramp fit on the non-destructive read-outs also considering
cosmic rays. From the beginning, pixels with data quality flags ``bad
detector pixels'', ``unstable response'', and ``bad
or uncertain flat value'' were masked out and excluded from
further analysis as suggested by previous transit exoplanet
spectroscopic observations\citep[e.g.][]{Berta2012, Kreidberg2014}.

One major challenge of high contrast imaging observation using WFC3/IR
is significant under-sampling of the detector.  2M1207 A and b are
only separated by $\sim6$ pixels or $\sim$5 FWHM of the PSF. When
applying roll subtraction, notable artifact structures are generated
by image interpolation and shifting. On the other hand, \tinytim{} PSF
simulator\citep{Krist1995} offers a solution by providing Nyquist or better
sampled PSF, but systematic errors of \tinytim{} PSF for WFC3 limits
its ability in high precision photometry\citep{Biretta2014}. However,
we are able to fully characterize the difference of model and observed
PSFs with 6 orbits time-resolved observation data. To obtain robust \tinytim{} PSF
photometry, we designed a 2-round PSF fitting strategy: 1. calculating
correction map for \tinytim{}; 2. hybrid PSF photometry.

For both of 2 rounds, we used {\em Tiny Tim} to calculate 10$\times$
over-sampled model PSFs based on the filters, the spectra 
\citep{Bonnefoy2014, Patience2010}, the telescope's actual focus, and
the telescope jitter.  We used the new set of Tiny Tim parameters
provided by \cite{Biretta2014} to improve model of the cold mask,
OTA spikes, and the coma. The focus parameters were calculated
using the model listed on the STScI
website\footnote{\url{http://www.stsci.edu/hst/observatory/focus/FocusModel}}. To
register \tinytim{} PSF to observed PSF of 2M1207A, we moved the
over-sampled PSF over a coordinate grid (gird size=0.001 pixel) using
cubic interpolation and minimize the rms difference of the
observed and re-binned \tinytim{} PSFs over a region centered on 2M1207A with a
5-pixel-radius aperture centered on 2M1207b excluded. Then we introduce
another \tinytim{} PSF for 2M1207b and  fit the
position of 2M1207 b and the photometries of 2M1207 A and b simultaneously
by least square optimization. From our data set, we discovered that
the difference of observed PSFs and model PSFs were very stable for
given PSF positions. Therefore, at the end of the first round PSF
fitting, we derived 8 (2 roll angles $\times$ 4 dithering positions)
observationally derived correction maps for each filter:
\begin{equation}
  \mathrm{Corr = Median(PSF_{obs.} - PSF_{model} )}
\end{equation}
where $\mathrm{PSF_{model}}$ is a combination of two \tinytim{} PSFs
for 2M1207 A and b. In the second round, we combined the correction
term linearly with the two \tinytim{} PSFs to generate hybrid PSFs,
and fitted the three components together. We found that by including the correction term,
the reduced $\chi^{2}$ of PSF fitting is decreased from $\sim 10$ to
around unity. Relative photometry is acquired from the scaling
parameters of the model PSFs.

PSF profiles change with exposure positions due to
pixelation, especially for the case that WFC3 IR is significantly
under-sampled. Also, the flat fields may potentially have large scale
structures \citep{dressel2012wide}. We found a correlation of
photometry with PSF positions on detector frame for both 2M1207 A and
b. Correction were made by normalizing each group of
exposures that have the same dithering position and roll angle
individually -- we took the median of the fluxes that were measured
from these exposures as normalization factors and divided them from
every photometric measurement. Because the normalization factor for
each group of exposures is calculated across the whole observation,
this normalization step has negligible impacts on variability
analysis.

@@

First we estimated the photon noises for the photometry of 2M1207b. We
propagated the photon noises of every single pixel that were
calculated from count rates and detector gain, to the PSF fitting
results. The photon noises for photometry in F125W and F160W are
1.33\% and 1.02\%, respectively.

Since the PSFs for the 2M1207 A and b were fitted simultaneously in the
PSF photometry procedure, the uncertainties of photometry and position
for the primary and secondary are coupled. Imperfection of position
measurement of 2M1207A could potentially affect the photometry of
2M1207b. We used a Monte Carlo (MC) method to
evaluate the overall systematic of the PSF fitting. We applied the
PSF photometry to images that were added with random Poisson noises and
repeat the photometry procedure for 1000 times. From the distribution of the
result, the uncertainty for F125W and F160W photometry are 
1.34\% and 1.12\%, respectively. We conclude that the white noise of
our observation is dominated by photon noise.

@@

2M1207b were observed at 8 different spots on the
detector (2 rolls $\times$ 4 dithering positions). Imperfect flat
field correction are potential sources of variation as they may
introduce position-dependent differences in the count rates. The
uncertainty of WFC3 IR pipeline flat field is typically $\sim 1\%$
\citep{dressel2012wide}.

In PSF photometry, however, multiple pixels are fitted simultaneously
and we expect a lower than 1\% uncertainty from the flat field
errors. To verify this, we multiplied every image by an artificial flat
field error mask (AFEM) -- a uniformly distributed Gaussian noise array with
mean of 1 and sigma of 1\% -- and repeated the PSF photometry on the
resulting images.  The analysis of these experiments showed almost
identical light curve to the original, verifying that the flat field
errors do not affect our photometry significantly (Figure
\ref{fig:2}).

\section{Verification of Photometric Variations and Amplitude
  Estimate}

@@

The light curves that resulted from our photometry show apparently
sinusoidal modulations, discussed in more details in
\S,\ref{Results}. To verify that these modulations are intrinsic to
the object and not result of our data reduction procedure or due to
instrumental changes, we carried out three different tests.

First, we fitted sine waves independently to the two filters to verify
the similarity of the signal in the two bands (Figure \ref{fig:2}). Inconsistent periods or
light curve shapes would argue against a genuine signal.  We found
that the periods of the best fit sine waves are similar,
$10.5^{+1.2}_{-1.3}$h for F125W and $9.1^{+1.1}_{-1.0}$h for
F160W. These periods are roughly consistent within the
uncertainty. Furthermore, these periods are not close to any
timescales over which HST or WFC3 are known to changes.

 As a second test we repeated the analysis neglecting the first
 orbit. The motivation behind this test was that, due to spacecraft
 thermal settling, the first orbit of HST observations is often
 slightly unstable and is often neglected in high-precision studies
 (\textbf{citation?}). Indeed, in our analysis 2M1207 A is
 significantly fainter in the first orbit (Figure \ref{fig:3}) than in
 the subsequent ones.
 Our analysis based on orbits 2--6 only found essentially identical
 results to our analysis of orbits 1--6, based on which we conclude
 that the first less reliable orbit does not affect our results
 significantly (Figure \ref{fig:2}, middle panel).

As a third test we explored whether a subset of images, perhaps
imperfectly normalized or correlated with specific instrument states,
could be driving the light curves into an apparently sinusoidal
shape. To test this possibility we split the data into two temporally
overlapping halves: sub-dataset one from dithering positions 1 and 3,
and sub-dataset two from dithering positions 2 and 4. For both
datasets we repeated our analysis independently.  For both of F125W
and F160W, two halves demonstrated similar trend of variability as
shown in Figure \ref{fig:2}.  Our analysis detected sinusoidal
modulations in {\em both} sub-datasets and in {\em both} filters, with
periods and amplitudes consistent with those derived from the complete
data set (\ref{fig:2}, upper panel).

 These tests demonstrate that the modulation seen in our data are
 consistently present in the different filters, in the different time
 segments of the data, and in data obtained in different dithering
 positions.

@@

To constrain the uncertainties of the least square optimization
results, we used a MC method to improve the fitting. We
generated a series of random Gaussian noises with the standard deviation same as
the photon noise, added them to the original light curves, and applied
least square fit to the new light curves. We repeated above routine for 100,000 times and obtained
the distribution of the fitting parameters. The distributions for the
periods and the amplitudes for F125W and F160W light curves are
shown in Figure \ref{fig:4}.

\section{Result}

\label{Results}
We present the first  high-resolution, high-cadence, and
high-precision photometry of a directly imaged planet or
planetary-mass companion. Our observations reveal a modulation in the
light curve of the $\sim 4 \mathrm{M_{{Jup}}}$ companion 2M1207b, the first detection
of modulations in directly imaged planetary mass objects.  The best fit
periods for F125W and F160W are 10.5 and 9.1 hour correspondingly. The
amplitudes for the normalized light curves are 1.45\% and 0.92\% for
F125W and F160W light curves.

We obtained high signal to noise photometry series for both 2M1207 A
and B (Figure \ref{fig:3}). On average, the photometric contrast is
$6.52\pm0.01$ mag for
F125W and $5.77\pm0.01$ mag for F160W. The difference of F125W contrast
from that measured in J-band \citep{Mohanty2007} and
F160W contrast from that measured with NICMOS F160W 
\citep{Song2006} is due to the different throughput profiles of the
filters.

The distributions for the periods demonstrate long tail shaped towards
long period, with core region roughly Gaussian. With 64\% confidence,
we estimated the 1-$\sigma$ range for the periods of F125W and F160W
to be $10.5_{-1.2}^{+1.3}$ and $9.1_{-1.0}^{+1.1}$ h,
respectively. The period of best fitted sine wave of F125W light
curve is 1.5h longer than that of F160W, and 0.3--0.4 h larger
1-$\sigma$ standard deviation. We also jointly fit the two band light
curves together forcing the periods of two sinusoids to be the
same. We derived a modulation period of $10.2^{+0.9}_{-0.8}$ h.

We discover that the variation amplitudes in the two bands are significantly
different. By fitting Gaussians to the MC fit result distributions, we
determined that variation amplitude of F125W is 1.45\% with a
standard deviation of 0.22\%, and that of F160W is 0.92\% with a
standard deviation of 0.20\%. The peaks of the two
histograms separate by more than 2-$\sigma$. The variation amplitude
for F125W is 1.58 times of that of F160W light curve.

>>>>

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage{color}
\bibliographystyle{apj}
\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=magenta]{hyperref}
\shortauthors{Zhou et al.}

@@

\title{Discovery of Rotational Modulations in the Planetary-mass
  Companion 2M1207b: A Slow Rotation Period and Cloud Height in a Low
  Gravity Atmosphere}
\shorttitle{Variability of 2M1207 b}
\author{Yifan Zhou\altaffilmark{1}, D\'aniel Apai\altaffilmark{1,2,3},
  Glenn Schneider\altaffilmark{1},  Mark S. Marley\altaffilmark{4}}

\altaffiltext{1}{Department of Astronomy/Steward Observatory, The
  University of Arizona, 933 N. Cherry Ave., Tucson, AZ, 85721, USA,
  \href{mailto:yifzhou@email.arizona.edu}{yifzhou@email.arizona.edu}
}
\altaffiltext{2}{Lunar and Planetary Laboratory, The University of
  Arizona, 1640 E. University Blvd., Tucson, AZ 85718, USA}
\altaffiltext{3}{Earths in Other Solar Systems Team, NASA Nexus for
  Exoplanet System Science}
\altaffiltext{4}{NASA Ames Research Center, Naval Air Station,
  Moffett Field,Mountain View, CA 94035, USA}

\begin{abstract}
  Rotational modulations of brown dwarfs have recently provided
  powerful constraints on the properties of ultra-cool atmospheres,
  including longitudinal and vertical cloud structures and cloud
  evolution. Furthermore, detection of periodic light curve  can directly probe the rotational periods of ultra-cool objects.

  We present here, for the first time, time-resolved high-precision
  photometric measurements of a planetary-mass companion, 2MASS1207b,
  to a brown dwarf primary. Using HST/WFC3 observation and point
  spread function fitting photometry with two spacecraft roll angles,
  we detect photometric modulations in the light curve. The amplitudes
  are 1.45\% in the F125W and 0.92\% in the F160W filters; we find a
  consistent period of $10.2^{+0.9}_{-0.8}$
  h and similar phase in both bands by fitting sinusoids to the light curves. {\color{red}The relative amplitudes in the two filters are very similar to
  that found in a recent study of a field (high-gravity) L-dwarf,
  suggesting that the cloud structures that introduce the photometric
  modulations are similar in high- and low-gravity
  objects.} Importantly, our study also measures, for the first time,
  the rotational period for directly an imaged planetary-mass companion.
\end{abstract}

\keywords{brown dwarfs -- planets and satellites: atmospheres -- planets
  and satellites: individual (2M1207 b) -- techniques: photometric}
\maketitle
%
\section{Introduction}

Presence of condensate clouds is one the most unique and essential
features of the ultra-cool atmosphere of direct imaged exoplanet and
brown dwarfs. The studies of formation and properties of condensate
clouds \citep[e.g.][]{Ackerman2001, Burrows2006a, Helling2008,
  Allard2012} have achieved great improvement in understanding the
cloud behavior across different spectral types, especially in the
explanation of L-T transition \citep[e.g.][]{Burrows2006a,
  Marley2010}.  Surface gravity is suggested to be another key
parameters in defining cloud structures besides spectral type. Low
surface gravity objects (e.g. HR8799 bcd, \cite{Marois2008a}, 2M1207
b, \cite{Chauvin2005}) show significantly redder color compared to
their spectral type matched brown dwarfs.  The anomalous colors of low
surface gravity objects are best explained by the existence of thick
clouds \citep{Barman2011, Skemer2011, Skemer2012}. However, due to
lack of observational constraint, the dependence of cloud properties
on surface gravity is not very well modeled.

Intensity modulations introduced by heterogeneous clouds can be
directly observed and studied via time resolved observation and
rotational mapping. These techniques isolate the effect of cloud
properties and obtained great success in determining the rotation
period and unveiling the structures of the atmosphere of brown dwarfs
\citep[e.g.][]{Apai2013,Buenzli2012,Buenzli2015,Burgasser2013,Radigan2012,Yang2014,Metchev2015}. \cite{Kostov2013}
demonstrated the great potential of these techniques in the study of
the cloud properties of direct imaged exoplanets and planetary mass
companions. However, high contrast magnifies
the challenges for direct imaged exoplanets and planetary mass
companions to acquire high-precision light curves comparing to brown dwarfs.

2M1207 b is a planetary mass companion that has mass of
$\sim 2.3-4.8M_{\mathrm{Jup}}$ \citep[][]{Barman2011} and an angular separation of $0.78''$ to its
host 2M1207 A, an M8 brown dwarf. Moderate contrast  \citep[6.7 mag
in J-band,][]{Mohanty2007} makes it one the most feasible
planetary mass companions for high-precision time-resolved observation.

@@

We obtained direct images of the 2M1207A+b system on UT 2014 April 11
from 08:07:47 to 16:53:18 using HST and its Wide Field Camera 3
\citep[WFC3, pixel scale=$\sim0.13''$, ][]{Kimble2008} in the frame of
the HST Proposal GO-13418 (PI: D. Apai). We acquired the observations
in filters F125W ($\lambda_{\mathrm{pivot}}$ = 1245.9 nm, full width
at half maximum (FWHM) = 301.5 nm) and F160W
($\lambda_{\mathrm{pivot}}$ 1540.52, FWHM = 287.9 nm), roughly
corresponding to the J and H bands. We used the $256\times256$ pixels
sub-array mode to avoid memory dumps during the observations.  In
order to provide a near-continuous coverage for detecting modulations
we observed the 2M1207 system in 6 consecutive HST orbits, obtaining
data with cadence of $\sim1.5$ minutes over a baseline of 8 hours and
40 minutes. The observations were interrupted by 58 minutes long Earth
occultations every 94 minutes.

The observations applied space craft rolls between each two orbits to
allow roll-subtraction of the primary \citep[e.g.][]{Song2006}. The
telescope roll angles for orbit 1, 3, and 5, and those for 2, 4, and
6 differ by $25^{\circ}$. At the separation of 2M1207 b, this angle
difference corresponds to a displacement of $0.34''$, or 2.75 and 2.30
resolution elements in F125W and F160W, respectively. In each orbit we
took 8 SPARS10 exposure sequence with NSAMP=10,
alternating between F160W and F125W filters, with 2--3 identical
exposures in one exposure sequence. To improve sampling and reduce the
risk that point spread function (PSF) lays on bad pixels, we applied a
4-point dither pattern with differential "X/Y" offsets of 1.375" in
the detector frame, providing optimal non-integral (half pixel)
stepping of 10.5 and 8.5 pixels in F125W and F160W, respectively. Over
the 6 orbits, we obtained 70 images with 10 non-destructive read-outs
in F125W and 64 images in F160W with exposure time of 88.4~s for both
filters.

@@

 \begin{figure*}
  \centering
  \plotone{systematics}
  \caption{F125W (left) and F160W (right) light curves under different
    variability verification tests. Individual measurements are
    plotted with gray crosses. Photometric measurements of the same exposure
    sequence are binned, and binned photometry are plotted with points
    or squares. Best fitted sinusoidal curves are plotted with solid
    lines. {\em Upper}: binned measurements taken in dithering
    position 1 and 3 (red points) and that taken in 2 and 4 (blue
    squares) are plotted with different symbols. They demonstrate same
    trend of modulation. In upper left panel, green line is sinusoidal
    wave fitted with all parameter set free, and purple line is
    sinusoidal wave fitted with period set the same as that of
    F125W. {\em Middle}: sinusoidal waves fitted without using the
    data taken in Orbit \#1. These curves are almost identical to the
    curves plotted in upper panel. {\em Lower}: photometry measured
    with AFEM-added images and best fitted sinul curves. These
    points and curves are also almost identical to those plotted
    in the upper panel.}
  \label{fig:2}
\end{figure*}

@@

We start the reduction from the \flt{} produced by the WFC3's
\texttt{calwfc3} pipeline. We do not opt to use \ima{} that contain
all non-destructive read-outs, because these provided less information
on 2M1207A, which saturated after the first few samples.  The \flt{}
are results of basic calibration, including dark current correction,
non-linearity correction, flat field correction, as well as
up-the-ramp fit on the non-destructive read-outs. From the beginning,
pixels with data quality flags ``bad detector pixels'', ``unstable
response'', and ``bad or uncertain flat value'' are masked out and
excluded from further analysis as suggested by previous transit
exoplanet spectroscopic observations\citep[e.g.][]{Berta2012,
  Kreidberg2014}.

One major challenge of high contrast imaging observation using WFC3/IR
is significant under-sampling of the detector.  2M1207 A and b are
only separated by $\sim6$ pixels or $\sim$5 FWHM of the PSF. When
applying roll subtraction, notable artifacts are generated
by image shifting and interpolation. On the other hand, \tinytim{} PSF
simulator\citep{Krist1995} offers a solution by providing Nyquist or better
sampled PSF, but systematic errors of \tinytim{} PSF for WFC3 limits
its ability in high precision photometry\citep{Biretta2014}. However,
we are able to fully characterize the difference of model and observed
PSFs with 6 orbits time-resolved observation data. To obtain robust \tinytim{} PSF
photometry, we design a 2-round PSF fitting strategy: 1. calculating
correction map for \tinytim{}; 2. hybrid PSF photometry.

For both of 2 rounds, we use \tinytim{} to calculate 10$\times$
over-sampled model PSFs based on the filters, the spectra
\citep{Bonnefoy2014, Patience2010}, the telescope's actual focus, and
the telescope jitter.  We use the set of \tinytim{} parameters
provided by \cite{Biretta2014} to improve modeling  the cold mask, OTA
spikes, and the coma. The focus parameters are calculated using the
model listed on the STScI
website\footnote{\url{http://www.stsci.edu/hst/observatory/focus/FocusModel}}.
To register the \tinytim{} PSF to the observed PSF of 2M1207 A, we move the
over-sampled PSF on a coordinate grid (gird size=0.001 pixel) using
cubic interpolation, and search for the position that minimizes the
rms difference of the observed and re-binned \tinytim{} PSFs over a
region centered on 2M1207 A with a 5-pixel-radius aperture centered on
2M1207 b excluded.  Then we introduce another \tinytim{} PSF for
2M1207 b and fit the position of 2M1207 b and the scales of the
/tinytim{} PSFs of 
2M1207 A and b simultaneously with least square optimization. From the
first round fit, we discover that the difference of observed PSFs and model
PSFs are very stable for given PSF positions. Therefore, at the end
of the first round PSF fitting, we derive 8 (2 roll angles $\times$ 4
dithering positions)  correction maps for each
filter:
\begin{equation}
  \mathrm{Corr = Median(PSF_{obs.} - PSF_{model} )}
\end{equation}
where $\mathrm{PSF_{model}}$ is a combination of two \tinytim{} PSFs
for 2M1207 A and b. In the second round, we combine the correction
term linearly with the two \tinytim{} PSFs to generate hybrid PSFs,
and fit the three components together. We find that by introducing the correction term,
the reduced $\chi^{2}$ of PSF fitting is decreased from $\sim 10$ to
$\sim 1$. Relative photometry is acquired from the scaling
parameters of the \tinytim{} PSFs.

PSF profiles change with exposure positions due to
pixelation, especially for the case that WFC3 IR is significantly
under-sampled. Also, the flat fields may potentially have large scale
structures \citep{dressel2012wide}. We find a correlation of
photometry with PSF positions on detector frame for both 2M1207 A and
b. Correction is made by normalizing each group of
exposures that have the same dithering position and telescope roll angle
individually -- we take the median of the fluxes that are measured
from these exposures as normalization factors and divided them from
every photometric measurement. Because the normalization factor for
each group of exposures is calculated across the whole observation in
time domain,
this step has negligible impacts on variability
analysis.

@@

First we estimated the photon noises for the photometry of
2M1207b. The total photon noise of the photometry is calculated by
combining the photon noise of every single pixel that is calculated
from count rates and detector gain. The photon noises for photometry
in F125W and F160W are 1.33\% and 1.02\%, respectively.

Since the PSFs for the 2M1207 A and b are fitted simultaneously, the
uncertainties of photometry and position for the primary and secondary
are coupled. Imperfection of position measurement of 2M1207 A could
potentially significantly affects the photometry of 2M1207 b. We used a Monte Carlo
(MC) method to evaluate the overall systematic of the PSF fitting. We
applied the PSF photometry to images that were added with random
Poisson noises and repeat the photometry procedure for 1000
times. From the distribution of the result, the uncertainties for F125W
and F160W photometry are 1.34\% and 1.12\%, respectively. We conclude
that the white noise of our observation is dominated by photon noise.

@@

2M1207 b were observed at 8 different spots on the
detector (2 rolls $\times$ 4 dithering positions). Imperfect flat
field correction can potentially introduce position-dependent differences in the count rates. The
uncertainty of WFC3 IR pipeline flat field is typically $\sim 1\%$
\citep{dressel2012wide}.

In PSF photometry, however, multiple pixels are fitted simultaneously
and we expect to be less affected by high spatial frequency flat field
noise, and a lower than 1\% uncertainty from the flat field
errors. To verify this, we multiplied every image by an artificial flat
field error mask (AFEM) -- a uniformly distributed Gaussian noise array with
mean of 1 and sigma of 1\% -- and repeated the PSF photometry on the
resulting images.  The analysis of these experiments showed almost
identical light curve to the original, verifying that the flat field
errors do not affect our photometry significantly (Figure
\ref{fig:2}, bottom panel).

\section{Verification of Photometric Modulations and Amplitude
  Estimate}

@@

The light curves that result from our photometry show apparently
sinusoidal modulations, discussed in more details in
\S,\ref{Results}. To verify that these modulations are intrinsic to
the object and not results of our data reduction procedures or due to
instrumental changes, we carry out three different tests.

First, we fit sinusoids independently to the light curves of two filters to verify
the similarity of the signal in the two bands (Figure \ref{fig:2}, top
panel). Inconsistent periods or
light curve shapes would argue against a genuine signal.  We find
that the periods of the best fit sine waves are similar,
$10.5^{+1.2}_{-1.3}$h for F125W and $9.1^{+1.1}_{-1.0}$h for
F160W. These periods are roughly consistent within the
uncertainty. Furthermore, these periods are not close to any
timescales over which HST or WFC3 are known to changes.

As a second test, we repeat the analysis neglecting the first
orbit. The motivation behind this test is that, due to spacecraft
thermal settling, the first orbits of HST observations are often
slightly unstable, and neglected in high-precision studies
\citep[e.g.][]{Mandell2013}. Indeed, in our analysis 2M1207 A is
significantly fainter in the first orbit (Figure \ref{fig:3}) than in
the subsequent ones.  Our analysis based on orbits 2--6 finds
essentially identical results to our analysis using the whole 6 orbits, based on
which we conclude that the first less reliable orbit does not affect
our results significantly (Figure \ref{fig:2}, middle panel).

As a third test, we explore whether a subset of images, perhaps due to
imperfectly normalized or correlated with specific instrument states,
can drive the light curves into an apparently sinusoidal shape. To
test this possibility, we split the data into two temporally
overlapping halves: subset one are images taken at dithering position
1 and 3, and subset two are those taken at dithering position 2 and
4. For both data-sets, we repeated our analysis independently.  For
both of F125W and F160W, two halves demonstrated similar trend of
variability.  Our analysis detect sinusoidal modulations in {\em
  both} subsets and in {\em both} filters, with periods and
amplitudes consistent with those derived from the complete data set
(Figure \ref{fig:2}, upper panel).

 These tests demonstrate that the modulation seen in our data are
 consistently present in the different filters, in the different time
 segments of the data, and in data obtained in different dithering
 positions. All of the three tests support the modulation signal to be
 intrinsic to the target.

@@

We use a MC method to analyze the light curve, and provide amplitudes
and periods as well as their uncertainties for both filters. We
generate series of random Gaussian noises with the standard deviation
same as the photon noises, add them to the original light curves, and
fit sinusoids to the newly combined light curves. We repeat above
routine for 100,000 times and obtain the distribution of the fitting
parameters (Figure \ref{fig:4}).

\section{Result}
\label{Results}

We present the first high-resolution, high-cadence, and high-precision
photometry of a directly imaged planet or planetary-mass
companion. Our observations reveal a modulation in the light curve of
the $\sim 4 \mathrm{M_{{Jup}}}$ companion 2M1207b, the first detection
of modulations in directly imaged planetary mass objects.  The best
fitted periods for F125W and F160W are 10.5 and 9.1 hour,
respectively. The modulation amplitudes for the normalized light
curves are 1.45\% and 0.92\% for F125W and F160W light curves.

We obtain high signal to noise photometry series for both 2M1207 A
and B (Figure \ref{fig:3}). On average, the photometric contrast is
$6.52\pm0.01$ mag for F125W and $5.77\pm0.01$ mag for F160W. The
difference of F125W contrast from that measured in J-band
\citep{Mohanty2007} and F160W contrast from that measured with NICMOS
F160W \citep{Song2006} is due to the different throughput profiles of
the filters.

The distributions for the periods demonstrate long tail shaped towards
long period, with core region roughly Gaussian. With 64\% confidence,
we estimate the 1-$\sigma$ range for the periods of F125W and F160W
to be $10.5_{-1.2}^{+1.3}$ and $9.1_{-1.0}^{+1.1}$ h,
respectively. The period of best fitted sinusoid for F125W light curve
is 1.5h longer than that for F160W, which is $\sim 20\%$ larger than
1-$\sigma$. We also jointly fit the two band light curves forcing the
periods of two sinusoids to be the same. We derive a modulation period of $10.2^{+0.9}_{-0.8}$ h.

We discover that the modulation amplitudes in the two bands are significantly
different. By fitting Gaussians to the MC fit result distributions, we
determine that modulation amplitude of F125W is 1.45\% with a
standard deviation of 0.22\%, and that of F160W is 0.92\% with a
standard deviation of 0.20\%. The peaks of the two
histograms separate by more than 2-$\sigma$. The modulation amplitude
for F125W is 1.58 times of that for F160W light curve.
